Presentation Title

### 2017-08-18 thesis/chapters/ch3-model.tex
Edit ch3

% As was shown in~\Cref{ch:problem}, the requirements are the following.
% \begin{itemize}
%     \item Global consensus
%     \item Security and fault tolerance
%     \item Performance and scalability
% \end{itemize}

% As mentioned in the Introduction, the primary goal is performance and scalability.
% Having a scalable blockchain system while still keeping global consensus
% allows the system to be ubiquitous and realise the full potential of blockchain.

% The secondary goal is to design an application neutral system.
% In particular, it should act as a backbone that provides the building blocks of blockchain based applications.
% It should be possible to buld any kind of application on top our system.
% Further, we do not impose on a consensus algorithm, as long as it satisfies the properties of atomic broadcast which we describe in~\Cref{sec:overview-cons}.

% The third and final goal is security.
% Our system should be unaffected in the presence of powerful adversaries.
% In particular, adversaries are Byzantine meaning that they can have arbitrary behaviour.
% Thus anything is possible from simply ommitting messages to colluding with eachother in order to undermine the whole system.

@@

\subsection{Extended TrustChain}
Extended TrustChain naturally builds on top of the standard TrustChain. 
Thus we first describe the standard TrustChain.
Our description has minor differences compared to the description in~\cite{multichain}.
This is to help with the description of the extended TrustChain.
We remark the difference when it occurs.
However, the two descriptions are functionally the same.

\subsubsection*{Standard TrustChain}
In TrustChain, every node has a ``personal'' chain. 
Initially, the chain only contains a genesis block generated by the nodes themselves.
When a node $A$ wishes to add a new transaction (TX) with $B$,
a new TX block is generated for both $A$ and $B$ and is appended to their respective chains.
The TX block must have a valid hash pointer pointing to the previous block
and a reference\footnote{This is different from the original TrustChain definition found in~\cite{multichain}.
In there, a TX block has two outgoing edges which are hash pointers to the two parties involved in the transaction.
This work uses one outgoing edge and a reference.} to its \emph{pair} on $B$'s chain.
An example of is shown in \Cref{fig:trustchain-bad}.

\begin{figure}
    \includegraphics[width=0.9\textwidth]{trustchain-bad}
    \centering
    \caption{Every block is denoted by $t_{i,j}$, where $i$ is the node ID and $j$ is the sequence number of the block.
    Thus we have three nodes and three corresponding chains in this example.
    The arrows represent hash pointers and the dotted lines represent references.
    The blocks at the ends of one dotted line are pairs of each other.
    The red block after $t_{b, 5}$ indicates a fork.}
    \label{fig:trustchain-bad}
\end{figure}

If every node follows the rules of TrustChain and we only consider hash pointers and not references,
then every chain effectively forms a singly linked list.
However, if a node violates the rules, then a \emph{fork} may happen.
That is, there may be more than one TX block with a hash pointer pointing back to the same block.
In \Cref{fig:trustchain-bad}, node $b$ (in the middle chain) created two TX blocks that both point to $t_{b, 5}$.
If this is a ledger system it can be seen as a double spend, where the currency accumulated up until $t_{b, 5}$ is spent twice.

\subsubsection*{Extended TrustChain}
We are now ready to explain the Extended TrustChain.
The primary difference is the introduction of a new type of block---checkpoint (CP) block.
In contrast to TX blocks, CP blocks do not store transactions or contain references.
Their purpose is to capture the state of the chain and the state of the whole system.
In particular, the state of the chain is captured with a hash pointer.
The state of the whole system is captured in the content of the CP block,
namely as a digest of the latest \emph{consensus result} which we explain in \Cref{sec:overview-cons}.
A visual representation is shown in \Cref{fig:trustchain-bad-cp}.

@@

\begin{figure}
    \includegraphics[width=0.9\textwidth]{trustchain-bad-cp}
    \centering
    \caption{The circles represent CP blocks,
    they also have hash pointers (arrow) but do not have references (dotted line).
    Note that the sequence number counter do not change, it is shared with TX blocks.}
    \label{fig:trustchain-bad-cp}
\end{figure}

\subsection{Consensus protocol}
\label{sec:overview-cons}

% The consensus protocol uses an existing Byzantine consensus algorithm (explained in detail in \Cref{sec:xyz}) and runs it continuously in rounds.
The consensus protocol can be seen as a technique of running infinitely many rounds of some 
Byzantine consensus algorithm\footnote{More accurately it is ACS or asynchronous subset consensus, we describe ACS in \Cref{sec:acs-background}.},
starting a new execution immediately after the previous one is completed.
This is necessary because blockchain systems always need to reach consensus on new values proposed by nodes in the system,
or CP blocks in our case.

The high communication complexity of Byzantine consensus algorithms prohibits us from running it on a large network.
Thus, for every round, we randomly select some node---called facilitators---to collect CP blocks from every other node and use them as the proposal.
The facilitators are elected using a \emph{luck value}, which is computed using $\textsf{H}(\C_r || pk_u)$,
where $\C_r$ is the consensus result 
(which can be seen as the set union of all the CP blocks collected by the facilitators)
in round $r$ and $pk_u$ is the public key of node $u$.
Intuitively, the election is guaranteed to be random 
because the output of a cryptographically secure hash function is unpredictable and $\C_r$ cannot be determined in advance.
Finally, the elected facilitators run a Byzantine consensus algorithm to agree on a set of CP blocks.
The next round begins when the new facilitators are computed from these CP blocks.

\subsection{Transaction and validation}
The TX protocol is a simple request and response protocol.
The nodes exchange one round of messages and create new TX blocks on their respective chains.
Thus, as we mentioned before, one transaction should result in two TX blocks.

The consensus and transaction protocol by themselves do not provide a mechanism to detect tampering.
Thus we need a validation protocol to counteract malicious behaviour.
When a node wishes to validate one of its TX, it asks the counterparty for the \emph{agreed fragment} of the TX.
An agreed fragment of a TX block is a section of the chain beginning and ending with CP blocks that contain the TX block,
the CP blocks must also be in consensus.
Upon the counterparty's response, the node checks that the CP blocks are indeed in consensus,
the hash pointers are valid and his TX is actually in the fragment.
The TX is valid if these conditions are satisfied.
Intuitively, this works because it is hard (because a cryptographically secure hash function is second-preimage resistant)
to create a different chain that begins and ends with the same two CP blocks.

\subsection{Combined protocol}
\label{sec:combined-protocol}
The final result is essentially the concurrent composition of the three aforementioned protocols,
all making use the Extended TrustChain data structure.

Our sub-protocol design gives us the highly desirable non-blocking property.
In particular, we do not need to ``freeze'' the state of the chain for some communication to complete in order to create a block.
For instance, a node may start the consensus protocol, and while it is running, create new transactions and validate old transactions.
By the time the consensus protocol is done, the new CP block is added to whatever the state that the chain is in.
It is not necessary to lock the chain while the consensus protocol is running and then unlock it afterwards.

@@

This section and the ones following it give a technical treatment of what the content in System overview.
For notational clarity, we use the following convention (adapted from~\cite{miller2016honey}) for most of this work.
\begin{itemize}
\item Lower case (e.g. $x$) denotes a scalar object or a tuple.
\item Upper case (e.g. $X$) denotes a set or a constant.
\item Sans serif (e.g. $\textsf{fn}(\cdot)$) denotes a function.
\item Monospace (e.g. $\texttt{ack}$) denotes message type.
\end{itemize}
Further, we use $a || b$ to denote concatenation of the binary representations of $a$ and $b$.

We assume purely asynchronous channels with eventual delivery.
Thus in no stage of our protocol do we make timing assumptions.
The adversary has full control of the delivery schedule and the message ordering of all messages.
But they are not allowed to drop messages except for their own\footnote{
    Reliability can be achieved in unreliable networks by resending messages or using some error correction code.
}.

@@

% TODO big N does not follow our convention
In our model, we consider $N$ nodes, which is the population size.
$n$ of them are facilitators, $t$ out of $n$ are malicious and the inequality
$n \ge 3t + 1$ must hold.
This is from the work of Pease, Shostak and Lamport, where they show a network of $n$ nodes cannot reach Byzantine agreement with $t \ge n/3$~\cite{pease1980reaching}.
Further, the inequality $N \ge n + t$ must also hold.
This is due to our system design, which becomes clear in~\Cref{sec:consensus-phase}.

Our threat model is as follows. 
We use a restricted version of the adaptive corruption model.
The first restriction is that corrupted node can only change across rounds.
That is, if a round has started, the corrupted nodes cannot change until the next round.
The second restriction is that the adversary, presumably controlling all the corrupted nodes, is forgetful.
Namely, the adversary may learn the internal state such as the private key of a corrupted node,
but if the node recovers, then the adversary must forget the private key.
This is realistic because otherwise the adversary can eventually learn all the private keys and sabotage the system.

\section{Extended TrustChain}
The primary data structure used in our system is the Extended TrustChain.
Each node $u$ has a public and private key pair---$pk_u$ and $sk_u$, and a chain $B_u$.
The chain consist of blocks $B_u = \{ b_{u, i} : i \in \{ 0, \dots, h - 1 \} \}$,
where $b_{u, i}$ is the $i$th block of $u$,
and $h$ is the height of the block (i.e. $h = |B_u|$).
We often use $b_{u, h}$ to denote the latest block.
There are two types of blocks, TX blocks and CP blocks.
If $T_u$ is the set of all TX blocks in $B_u$ and $C_u$ is the set of all CP blocks is $B_u$,
then it must be the case that $T_u \cup C_u = B_u$ and $T_u \cap C_u = \varnothing$.
The notation $b_{u, i}$ is generic over the block type.
We assume there exists a function $\textsf{typeof}: B_u \rightarrow \{ \tau, \gamma \}$ that returns the type of the block,
where $\tau$ represents the TX type and $\gamma$ represents the CP type.

@@

Our consensus protocol runs in rounds as discussed in \Cref{sec:system-overview}.
Every round is identified by a round number $r$, which is incremented on every new round.
The consensus result is a tuple, i.e. 
$$\C_r = \langle r, C \rangle,$$
where $C$ is a set of CP blocks agreed by the facilitators of round $r$.
\end{definition}

@@

If there exist a tuple $\langle c_{u,a}, c_{u, b} \rangle$ for a TX block $t_{u, i}$,
where 
$$a = \argmin_{k, k < i, \textsf{typeof}(b_{u,k}) = \gamma}(i - k)$$
$$b = \argmin_{k, k > i, \textsf{typeof}(b_{u,k}) = \gamma}(k - i),$$
then $\langle c_{u,a}, c_{u, b} \rangle$ is the enclosure of $t_{u, i}$.
Intuitively, that is the two closest CP blocks that surround $t_{u, i}$.
Note that $c_{u, a}$ is the more recent CP block.
Also, some TX blocks may not have any enclosure, then its enclosure is $\bot$.
Agreed enclosure is the same as enclosure with an extra constraint where the CP blocks must be in some consensus result $\C_r$.

@@

Note that the ``closest CP blocks'' property must also apply to agreed enclosure. 
Suppose a chain is in the form
$\{c_{i}, c_{i+1}, t_{i+2}, c_{i+3}\}$\footnote{Usually the notation is of the form $c_{u, i}$, but the node identity is not important here so we simplify it to $c_{i}$}
and $c_{i}, c_{i+1}, c_{i+3}$ are in $\C_r, \C_{r+1}, \C_{r+3}$ respectively,
then the agreed enclosure of $t_{i+2}$ is $\langle c_{i+1}, c_{i+3}\rangle$ and cannot be $\langle c_{i}, c_{i+3}\rangle$.

@@

Our scalable consensus protocol runs on top of Extended TrustChain.
It uses an unmodified asynchronous subset consensus (ACS) algorithm as the key building block.
The objectives of the protocol are to
    allow honest nodes always make progress (in the form of creating new CP blocks),
    compute correct consensus result in every round
    and have an unbiased election of facilitators.
Concretely, we define the necessary properties as follows.
\begin{definition}
\label{def:consensus}
\textbf{\emph{Properties of the consensus protocol}}

$\forall r \in \mathbb{N}$, the following properties must hold.
\begin{enumerate}
    \item \emph{Agreement}:
        If one correct node outputs a set of facilitators $\F_r$,
        then every node outputs $\F_r$
    \item \emph{Validity}:
        If any correct node outputs $\F_r$, then 
            \begin{enumerate}
                \item $|\C_r| \ge N - t$ and each CP in $\C_r$ belong to a different node\footnote{
                $\C_r$ is a tuple but we abuse the notation here by writing $|\C_r|$ to mean the number of CP blocks in the second element of $\C_r$.}.
                \item $\F_r$ must contain at least $n - t$ honest nodes and
                \item $|\F_r| = n$.
            \end{enumerate}
    \item \emph{Fairness}:
        Every node with a CP block in $\C_r$ should have an equal probability of becoming a member of $\F_r$.
    \item \emph{Termination}:
        Every correct node eventually outputs some $\F_r$.
\end{enumerate}
\end{definition}
These properties are similar to Byzantine consensus properties (which we describe next in \Cref{sec:acs-background}), but there are subtle differences.
Firstly, they are properties for every node in the network and not just the facilitators.
Secondly, they must be satisfied for all rounds,
because a failure in one round will lead to failures in sebsequent rounds.

Before describing the protocol in detail,
we take a brief detour to give background on the ACS protocol as it is the primary building block.
Then we move on to describe the two phases of our consensus protocol---bootstrap phase and consensus phase.

@@

There are $n$ nodes, of which at most $t$ might experience Byzantine fault.
Node $i$ starts with an input value $v_i$.
The nodes must decide for one of  those values, satisfying the following.
\begin{enumerate}
    \item \emph{Agreement:}
        If a correct node outputs $v$, then every node outputs $v$.
    \item \emph{Validity:}
        The decision value must be the input value of a node.
    \item \emph{Termination:}
        All correct nodes terminate in finite time.
\end{enumerate}
\end{definition}
A node under Byzantine fault means that it can have arbitrary behaviour.
For example not sending a message or colluding with other Byzantine nodes to undermine the entire system.
Note that the decision is on a single value.
This is in contrast to ACS which we describe next.

@@

ACS has the nice property of censorship resilience when compared to other consensus algorithms.
For instance, Hyperledger and Tendermint uses a variant of Practical Byzantine Fault Tolerance (PBFT)~\cite{castro1999practical} as their consensus algorithm.
In PBFT, a leader is elected, if the leader is malicious but follows the protocol, then it can selectively filter transactions.
In contrast, every party in ACS are involved in the proposal phase,
and it guarantees that if $n - 2t$ parties propose the same transaction, then it must be in the agreed output.
Thus, if some value is submitted to at least $n - 2t$ nodes, it is guaranteed to be in the consensus result.
For a detailed description of ACS, we refer to the HoneyBadgerBFT work~\cite{miller2016honey}.
To understand the remainder of this work, the knowledge of~\Cref{def:acs} is sufficient.

@@

To bootstrap, imagine that there is some bootstrap oracle, that initiates the code on every node.
The code satisfies all the properties in \Cref{def:consensus}.
Namely, every node has the same set of valid facilitators $\F_1$ that are randomly chosen.
This concludes the bootstrap phase.
For any future rounds, the consensus phase is used.

@@

\subsection{Consensus phase}
\label{sec:consensus-phase}
The consensus phase begins when $\F_r$ is available to all the nodes.
Note that $\F_r$ indicates the facilitators that were elected using results of round $r$ and are responsible for driving the ACS protocol in round $r + 1$.
The goal is to reach agreement on a set of new facilitators $\F_{r + 1}$ that satisfies the four properties in \Cref{def:consensus}.

There are two scenarios in the consensus phase.
First, if node $u$ is not the facilitator, it sends $\langle \texttt{cp\_msg}, c_{u, h} \rangle$ to all the facilitators.
Second, if the node is a facilitator, it waits until $N - t$ messages of type \texttt{cp\_msg} are received.
% Second if the node is a facilitator, it waits for some duration $D$ and collect messages of type \texttt{cp\_msg}, where $D \gg \Delta$.
Invalid messages are removed.
That is blocks with invalid signatures and blocks signed by the same key.
With the sufficient number of \texttt{cp\_msg} messages,
it begins the ACS algorithm and some $\C'_{r + 1}$ should be agreed upon by the end of it.
Duplicates and blocks with invalid signatures are again removed from $\C'_{r+1}$ and we call the final result $\C_{r+1}$.
We have to remove invalid blocks a second time (after ACS) because the adversary may send different CP blocks to different facilitators,
which results in invalid blocks in the ACS output, but not in any of the inputs.

The core of the consensus phase is the ACS protocol.
While any ACS protocol that satisfies the standard definition will work,
we use a simplification of HoneyBadgerBFT~\cite{miller2016honey} as our ACS protocol
because it is a consensus algorithm designed for blockchain systems.
We do not use the full HoneyBadgerBFT due to the following.
First, the transactions in HoneyBadgerBFT are first queued in a buffer and the main consensus algorithm starts only when the buffer reaches an optimal size.
We do not have an infinite stream of CP blocks, thus buffering is unsuitable.
Second, HoneyBadgerBFT uses threshold encryption to hide the content of the transactions.
But we do not reach consensus on transactions, only CP blocks, so hiding CP blocks is meaningless for us as it contains no transaction information.

Continuing, when $\F_{r}$ reaches agreement on $\C_{r+1}$,
they immediately broadcast two messages to all the nodes---
first the consensus message $\langle \texttt{cons\_msg}, \C_{r+1} \rangle$,
and second the signature message $\langle \texttt{cons\_sig}, r, sig \rangle$.
The reason for sending \texttt{cons\_sig} is the following.
Recall that channels are not authenticated, 
and there are no signatures in $\C_{r + 1}$.
If a non-facilitator sees some $\C_{r + 1}$, it cannot immediately trust it because it may have been forged.
Thus, To guarantee authenticity, every facilitator sends an additional message that is the signature of $\C_{r + 1}$.

@@

\subsection{Validity definition}
A transaction can be in one of three states in terms of validity---\emph{valid}, \emph{invalid} and \emph{unknown}.
Given a fragment $F_{v, j}$, the validity of the TX block $t_{u, i}$ with its corresponding fragment $F_{u, i}$ is captured by the function $\textsf{get\_validity}(t_{u, i}, F_{u, i}, F_{v, j})$ (\Cref{alg:get-validity}).
Note that $t_{u, i}$ and $F_{u, i}$ are assumed to be valid,
otherwise the node calling the function would have no point of reference.
This is not difficult to achieve,
because typically the caller is $u$, so it knows its own TX block and the corresponding agreed fragment.
If the caller is not $u$, it can always query for the agreed fragment that contains the transaction of interest from $u$.

\Cref{alg:get-validity} works as follows.
Before~\Cref{line:valid-fragment},
we essentially check whether the fragment is the one that the verifier needs.
If it is not, then the verifier cannot make any decision and return \emph{unknown}.
This is likely to be the case for new transactions because the result of $\textsf{agreed\_fragment}(\cdot)$ would be $\bot$.
The next two conditions check for tampering or missing blocks, if any of these misconducts are detected, then the TX is invalid.

>>>>

@@

\subsection*{TrustChain and Extended TrustChain}
Extended TrustChain naturally builds on top of the standard TrustChain. 
Thus we first describe the standard TrustChain.
Our description has minor differences compared to the description in~\cite{multichain},
which is to help with the description of the Extended TrustChain.
We remark the differences when they occur.

% \subsubsection*{Standard TrustChain}
In TrustChain, every node has a ``personal'' chain. 
Initially, the chain only contains a genesis block generated by the nodes themselves.
When a node $A$ wishes to add a new transaction (TX) with $B$,
a new TX block is generated for both $A$ and $B$ and is appended to their respective chains.
The TX block must have a valid hash pointer pointing to the previous block and a reference\footnote{This definition is different from the original TrustChain definition found in~\cite{multichain}.
    In the original definition, a TX block has two outgoing edges which are hash pointers to the two parties involved in the transaction.
    This work uses one outgoing edge and a reference.} to its \emph{pair} on $B$'s chain.
An example of is shown in \Cref{fig:trustchain-bad}.

\begin{figure}
    \includegraphics[width=0.9\textwidth]{trustchain-bad}
    \centering
    \caption{
    A TrustChain example with three nodes.
    Every block is denoted by $t_{i,j}$, where $i$ is the node ID and $j$ is the sequence number of the block.
    The arrows represent hash pointers and the dotted lines represent references.
    The blocks at the ends of one dotted line are pairs of each other.
    The red block after $t_{b, 5}$ indicates a fork.}
    \label{fig:trustchain-bad}
\end{figure}

If every node follows the rules of TrustChain and we only consider hash pointers and not references,
then every chain effectively forms a singly linked list.
However, if a node violates the rules,
either by accident or for malicious intent,
then a \emph{fork} may happen.
That is, there may be more than one TX block with a hash pointer pointing back to the same block.
In \Cref{fig:trustchain-bad}, node $b$ (in the middle chain) created two TX blocks that both point to $t_{b, 5}$.
If this is a ledger system it can be seen as a double spend, where the currency accumulated up until $t_{b, 5}$ is spent twice.

% \subsubsection*{Extended TrustChain}
The primary difference is the introduction of a new type of block---checkpoint (CP) block.
The most important function of the CP block is to represent the state of a node, in the form of a hash pointer.
A collection of CP blocks from all nodes represents the state of the whole system.
Therefore, nodes need to reach consensus on CP blocks using some consensus protocol, which we explain next.
A visual representation is shown in \Cref{fig:trustchain-bad-cp}.

@@

\begin{figure}
    \includegraphics[width=0.9\textwidth]{trustchain-bad-cp}
    \centering
    \caption{
    An example of Extended TrustChain.
    The circles represent CP blocks,
    they also have hash pointers (arrow) but do not have references (dotted line).
    Note that the sequence number counter do not change, it is shared with TX blocks.}
    \label{fig:trustchain-bad-cp}
\end{figure}

\subsection*{Consensus protocol}

The consensus protocol can be seen as a technique of running infinitely many times of some 
Byzantine consensus algorithm\footnote{More accurately it is asynchronous common subset (ACS) which we describe in \Cref{sec:acs-background}.},
starting a new execution immediately after the previous one is completed.
This approach is necessary because blockchain systems always need to reach consensus on new values proposed by nodes in the system,
or CP blocks in our case.

The high communication complexity of Byzantine consensus algorithms prohibits us from running it on a large network.
Thus, at the beginning of every Byzantine consensus algorithm execution,
we randomly elect a set of nodes---called facilitators---to collect CP blocks from every other node and use them as the input to the Byzantine consensus algorithm, and then run it.
After running the algorithm,
the facilitators output a set of CP blocks which we call the consensus result.
The result is then propagated to the network and the next algorithm execution begins.

\subsection*{Transaction and validation}
The TX protocol is a simple request and response protocol.
The nodes exchange one round of messages and create new TX blocks on their respective chains.
Thus, as we mentioned before, one transaction should result in two TX blocks.

The consensus and transaction protocol by themselves do not provide a mechanism to detect malicious behaviour such as tampering the TrustChain.
Thus, we need a validation protocol to counteract such behaviour.
When a node wishes to validate one of its transactions, it asks the counterparty for the \emph{agreed fragment} of the transaction.
Which is a fragment of the counterparty's chain beginning and ending with CP blocks but contains the TX block belonging to that transaction,
the CP blocks must also be in consensus.
Upon the counterparty's response, the node checks whether the CP blocks are in some consensus result,
the hash pointers are valid and his TX is actually in the fragment.
The TX is valid if these conditions are satisfied.
Intuitively, this works because it is hard (because a cryptographically secure hash function is collision resistant)
to create a different chain that begins and ends with the same two CP blocks.

\subsection*{Complete protocol}
The complete protocol is essentially the concurrent composition of the three aforementioned protocols,
all making use the Extended TrustChain data structure.

Our sub-protocol design gives us the highly desirable non-blocking property.
In particular, we do not need to ``freeze'' the state of the chain for some communication to complete in order to create a block.
For instance, a node may start the consensus protocol, and while it is running, create new transactions and validate old transactions.
By the time the consensus protocol is done, the new CP block is added to whatever state the chain is in.
It is not necessary to lock the chain while the consensus protocol is running and then unlock it afterwards.

@@

This section and the ones following it give a technical treatment of what the content in System overview.
For notational clarity, we use the following convention (adapted from~\cite{miller2016honey}) for most of this work.
\begin{itemize}
\item Lower case (e.g., $x$) denotes a scalar object or a tuple.
\item Upper case (e.g., $X$) denotes a set or a constant.
\item Sans serif (e.g., $\textsf{fn}(\cdot)$) denotes a function.
\item Monospace (e.g., $\texttt{ack}$) denotes message type.
\end{itemize}
Further, we use $a || b$ to denote the concatenation of the binary representations of $a$ and $b$.

We assume purely asynchronous channels with eventual delivery.
Thus in no stage of our protocol do we make timing assumptions.
The adversary has full control of the delivery schedule and the message ordering of all messages.
% But they are not allowed to drop messages except for their own\footnote{
%     Reliability can be achieved in unreliable networks by resending messages or using some error correction code.
% }.

@@

% TODO big N does not follow our convention
In our model, we consider $N$ nodes, which is the population size.
$n$ of them are facilitators, $t$ out of $n$ are malicious and the inequality
$n \ge 3t + 1$ must hold.
This inequality is from the work of Pease, Shostak and Lamport, where they show a network of $n$ nodes cannot reach Byzantine agreement with $t \ge n/3$~\cite{pease1980reaching}.
Further, the inequality $N \ge n + t$ must also hold so that there are enough honest nodes in the network,
the exact reason becomes clear in~\Cref{sec:consensus-phase}.

Our threat model is as follows. 
We use a restricted version of the adaptive corruption model.
The first restriction is that corrupted node can only change across rounds.
That is, if a round has started, the corrupted nodes cannot change until the next round.
The second restriction is that the adversary, presumably controlling all the corrupted nodes, is forgetful.
Namely, the adversary may learn the internal state such as the private key of a corrupted node,
but if the node recovers, then the adversary must forget the private key.
This restriction is realistic because otherwise the adversary can eventually learn all the private keys and sabotage the system.

\section{Extended TrustChain}
The primary data structure used in our system is the Extended TrustChain.
Each node $u$ has a public and private key pair---$pk_u$ and $sk_u$, and a chain $B_u$.
The chain consist of blocks $B_u = \{ b_{u, i} : i \in \{ 0, \dots, h - 1 \} \}$,
where $b_{u, i}$ is the $i$th block of $u$,
and $h$ is the height of the block (i.e. $h = |B_u|$).
We often use $b_{u, h}$ to denote the latest block.
There are two types of blocks, TX blocks and CP blocks.
If $T_u$ is the set of all TX blocks in $B_u$ and $C_u$ is the set of all CP blocks is $B_u$,
then $T_u \cup C_u = B_u$ and $T_u \cap C_u = \varnothing$.
The notation $b_{u, i}$ is generic over the block type.
% We assume there exists a function $\textsf{typeof}: B_u \rightarrow \{ \tau, \gamma \}$ that returns the type of the block,
% where $\tau$ represents the TX type and $\gamma$ represents the CP type.

@@

Our consensus protocol runs in rounds,
where the first round is defined to be 1 and it is incremented after every execution of the consensus protocol.
The consensus result is a tuple, i.e. 
$$\C_r = \langle r, C \rangle,$$
where $C$ is a set of CP blocks agreed by the facilitators of round $r$.
\end{definition}

@@

If there exist a tuple $\langle c_{u,a}, c_{u, b} \rangle$ for a TX block $t_{u, i}$, where
\begin{itemize}
\item $c_{u, a}$ is the closest CP block to $t_{u, i}$ with a lower sequence number and
\item $c_{u, b}$ is the closest CP block to $t_{u, i}$ with a higher sequence number,
\end{itemize}
then $\langle c_{u,a}, c_{u, b} \rangle$ is the enclosure of $t_{u, i}$.
Some TX blocks may not have any enclosure, then their enclosure is $\bot$.
Agreed enclosure is the same as enclosure with an extra constraint where the CP blocks must be in some consensus result $\C_r$.

@@

% Note that the ``closest CP blocks'' property must also apply to agreed enclosure. 
% Suppose a chain is in the form
% $\{c_{i}, c_{i+1}, t_{i+2}, c_{i+3}\}$\footnote{Usually the notation is of the form $c_{u, i}$, but the node identity is not important here so we simplify it to $c_{i}$}
% and $c_{i}, c_{i+1}, c_{i+3}$ are in $\C_r, \C_{r+1}, \C_{r+3}$ respectively,
% then the agreed enclosure of $t_{i+2}$ is $\langle c_{i+1}, c_{i+3}\rangle$ and cannot be $\langle c_{i}, c_{i+3}\rangle$.

@@

Our consensus protocol runs on top of Extended TrustChain.
It uses an unmodified asynchronous common subset (ACS) algorithm as the key building block.
The objectives of the protocol are to
    allow honest nodes always make progress (in the form of creating new CP blocks),
    compute correct consensus result in every round
    and have an unbiased election of facilitators.
Concretely, we define the following.
\begin{definition}
\label{def:consensus}
\textbf{\emph{TrustChain consensus}}

A TrustChain consensus protocol is correct if the following holds for every round $r$:
\begin{enumerate}
    \item \emph{Agreement}:
        If one correct node outputs a set of facilitators $\F_r$,
        then every node outputs $\F_r$
    \item \emph{Validity}:
        If any correct node outputs $\F_r$, then 
            \begin{enumerate}
                \item $|\C_r| \ge N - t$ and each CP in $\C_r$ belong to a different node\footnote{
                $\C_r$ is a tuple but we abuse the notation here by writing $|\C_r|$ to mean the number of CP blocks in the second element of $\C_r$.}.
                \item $|\F_r| = n$.
            \end{enumerate}
    \item \emph{Fairness}:
        Every node with a CP block in $\C_r$ should have an equal probability of becoming a member of $\F_r$.
    \item \emph{Termination}:
        Every correct node eventually outputs some $\F_r$.
\end{enumerate}
\end{definition}
These properties are similar to Byzantine consensus properties (which we describe next in \Cref{sec:acs-background}), but there are subtle differences.
Firstly, they are properties for every node in the network and not just the facilitators.
Secondly, they must be satisfied for all rounds because a failure in one round will lead to failures in subsequent rounds.

Before describing the protocol in detail,
we take a brief detour to give background on ACS as it is the primary building block.
Then we move on to describe the two phases of our consensus protocol---bootstrap phase and consensus phase.

@@

There are $n$ nodes, of which at most $t$ might experience Byzantine fault.
Node $i$ starts with an input value $v_i$.
The nodes must decide for one of  those values, satisfying the following.
\begin{enumerate}
    \item \emph{Agreement:}
        If a correct node outputs $v$, then every correct node outputs $v$.
    \item \emph{Validity:}
        The decision value must be the input value of a correct node.
    \item \emph{Termination:}
        All correct nodes terminate in finite time.
\end{enumerate}
\end{definition}
A node under Byzantine fault means that it can have arbitrary behaviour,
for example, not sending a message or colluding with other Byzantine nodes to undermine the entire system.
Note that the decision is on a single value,
which is in contrast to ACS that we describe next.

@@

ACS has the nice property of censorship resilience when compared to other consensus algorithms.
For instance, Hyperledger and Tendermint use a variant of Practical Byzantine Fault Tolerance (PBFT)~\cite{castro1999practical} as their consensus algorithm.
In PBFT, a leader is elected, if the leader is malicious but follows the protocol, then it can selectively filter transactions.
In contrast, every party in ACS are involved in the proposal phase,
and it guarantees that if $n - 2t$ parties propose the same transaction, then it must be in the agreed output.
Thus, if some value is submitted to at least $n - 2t$ nodes, it is guaranteed to be in the consensus result.
For a detailed description of ACS, we refer to the HoneyBadgerBFT work~\cite{miller2016honey}.
To understand the remainder of this work, the knowledge of~\Cref{def:acs} is sufficient.

@@

To bootstrap, imagine that there is some bootstrap oracle that initiates the correct program on every node,
meaning that it satisfied the properties in \Cref{def:consensus}.
This concludes the bootstrap phase.
For any future rounds, the consensus phase is used.

@@

\subsection{Consensus phase}
\label{sec:consensus-phase}
The consensus phase begins when $\F_r$ is available to all the nodes.
Note that $\F_r$ indicates the facilitators that were elected using results of round $r$ and are responsible for driving the ACS algorithm in round $r + 1$.
The goal is to reach agreement on a set of new facilitators $\F_{r + 1}$ that satisfies the four properties in \Cref{def:consensus}.

There are two scenarios in the consensus phase.
First, if node $u$ is not the facilitator, it sends $\langle \texttt{cp\_msg}, c_{u, h} \rangle$ to all the facilitators,
where $c_{u, h}$ is the most recent CP block.
Second, if the node is a facilitator, it waits until $N - t$ messages of type \texttt{cp\_msg} are received.
% Second if the node is a facilitator, it waits for some duration $D$ and collect messages of type \texttt{cp\_msg}, where $D \gg \Delta$.
Invalid messages are removed.
That is blocks with invalid signatures and blocks signed by the same key.
With the sufficient number of \texttt{cp\_msg} messages,
it begins the ACS algorithm and some $\C'_{r + 1}$ should be agreed upon by the end of it.
Duplicates and blocks with invalid signatures are again removed from $\C'_{r+1}$ and we call the final result $\C_{r+1}$.
We have to remove invalid blocks a second time (after ACS) because the adversary may send different CP blocks to different facilitators,
which results in invalid blocks in the ACS output, but not in any of the inputs.

The core of the consensus phase is the ACS algorithm.
While any ACS algorithm that satisfies the standard definition will work,
we use a simplification of HoneyBadgerBFT~\cite{miller2016honey} as our ACS algorithm 
because it is a consensus algorithm designed for blockchain systems.
We do not use the full HoneyBadgerBFT due to the following.
First, the transactions in HoneyBadgerBFT are first queued in a buffer and the main consensus algorithm starts only when the buffer reaches an optimal size.
We do not have an infinite stream of CP blocks, thus buffering is unsuitable.
Second, HoneyBadgerBFT uses threshold encryption to hide the content of the transactions.
But we do not reach consensus on transactions, only CP blocks, so hiding CP blocks is meaningless for us as it contains no transaction information.

Continuing, when $\F_{r}$ finish the ACS execution and reach agreement on $\C_{r+1}$,
they immediately broadcast two messages to all the nodes---
first the consensus message $\langle \texttt{cons\_msg}, \C_{r+1} \rangle$,
and second the signature message $\langle \texttt{cons\_sig}, r, sig \rangle$.
The reason for sending \texttt{cons\_sig} is the following.
Recall that channels are not authenticated, 
and there are no signatures in $\C_{r + 1}$.
If a non-facilitator sees some $\C_{r + 1}$, it cannot immediately trust it because it may have been forged.
Thus, To guarantee authenticity, every facilitator sends an additional message that is the signature of $\C_{r + 1}$.

@@

\subsection{Validity definition}
A transaction can be in one of three states in terms of validity---\emph{valid}, \emph{invalid} and \emph{unknown}.
Given a fragment $F_{v, j}$, the validity of the TX block $t_{u, i}$ with its corresponding fragment $F_{u, i}$ is captured by the function $\textsf{get\_validity}(t_{u, i}, F_{u, i}, F_{v, j})$ (\Cref{alg:get-validity}).
Note that $t_{u, i}$ and $F_{u, i}$ are assumed to be valid,
otherwise the node calling the function would have no point of reference.
This assumption is reasonable in practice because typically the caller is $u$,
so it knows its own TX block and the corresponding agreed fragment.
If the caller is not $u$, it can always query for the agreed fragment that contains the transaction of interest from $u$.

\Cref{alg:get-validity} works as follows.
Before~\Cref{line:valid-fragment},
we essentially check whether the fragment is the one that the verifier needs.
If it is not, then the verifier cannot make any decision and return \emph{unknown},
which is likely to be the case for new transactions because the result of $\textsf{agreed\_fragment}(\cdot)$ would be $\bot$.
The next two conditions check for tampering or missing blocks, if any of these misconducts are detected, then the TX is invalid.
